Analyzing Crime Trends in Toronto: Insights from the Police Annual Statistical Report

### 2024-09-27 paper/paper.qmd
Minor change.

# Introduction
Crime and public safety are critical concerns for urban environments like Toronto. Understanding the nature and distribution of crimes helps policymakers and law enforcement agencies allocate resources and develop effective crime prevention strategies. This paper examines Toronto’s crime patterns over the past decade, using the Toronto Police Annual Statistical Report to evaluate different types of crimes across the city's divisions [@OpenDataDataset].

Our analysis covers from 2014 to 2023 and focuses on two main crime categories: Crimes Against Property and Crimes Against the Person. We explore how incidents vary by division and examine clearance rates (resolved cases) to evaluate the effectiveness of law enforcement strategies. By highlighting regional crime trends, this paper provides insights into public safety concerns that can shape future policies.

>>>>

# Introduction
Crime and public safety are critical concerns for urban environments like Toronto. With a growing population and increasing urbanization, it is vital to understand the nature and distribution of crimes. Crime statistics offer insights into the effectiveness of policing efforts, trends in specific types of crimes, and the socio-economic factors that may influence these trends. Despite the availability of crime data through public portals like Open Data Toronto, gaps remain in how this data is analyzed and utilized to address public safety concerns [@OpenDataDataset]. Additionally, the dynamic nature of urban crime necessitates an ongoing, detailed analysis to ensure that resource allocations remain effective and that public safety strategies are continually optimized.

This paper focuses on analyzing crime trends in Toronto from 2014 to 2023 using the Toronto Police Annual Statistical Report. Specifically, it examines two broad categories of crime: Crimes Against Property (such as auto theft and break-ins) and Crimes Against the Person (such as assault). The primary objective of this analysis is to evaluate how these crimes are distributed across Toronto's police divisions and how effectively law enforcement is addressing these incidents, measured by clearance rates (the percentage of resolved cases).

A clear gap exists in understanding how different divisions are affected by varying types of crimes and how effectively these crimes are being addressed. This analysis aims to fill that gap by offering a division-level examination of crime trends, using detailed statistical tools to identify key areas where law enforcement resources may be optimized. The paper also explores the relationship between crime types and clearance rates, highlighting divisions where law enforcement may struggle to address specific types of crimes, such as property offenses, which tend to have lower resolution rates.

In conducting this analysis, data from the Open Data Toronto portal was cleaned, standardized, and analyzed using R, revealing significant differences in crime trends between Toronto's police divisions [@citeR]. The findings indicate that certain divisions, particularly those in urban centers, report higher incidences of property crimes, while more suburban divisions face a higher prevalence of violent crimes. Moreover, clearance rates for property crimes, such as auto theft, remain low, indicating a need for enhanced crime prevention and resolution strategies in these areas. These insights are critical for shaping future public safety policies and allocating law enforcement resources more effectively.
